---
title: "Week1Assignment DATA607"
author: "Banu Boopalan"
date: "8/31/2019"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
  ioslides_presentation: default
  pdf_document: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(tidyverse)  ## loads many but not all "tidyverse" packages
```

## Section 1: Read in DataSet and Report  Summary

```{r}
#Read with factors - Stringasfactors = TRUE to show summary counts
newdata <- read.csv(file="C:/Users/Banu/Documents/RScriptfiles/New folder/agaricus-lepiota.csv", header=FALSE, sep=",")
names(newdata) <- c("classes","cap-shape:","cap-surface:","cap-color:","bruises:","odor:","gill-attachment:","gill-spacing:","gill-size:","gill-color:","stalk-shape:","stalk-root:","stalk-surface-above-ring:","stalk-surface-below-ring:","stalk-color-above-ring:","stalk-color-below-ring:","veil-type:","veil-color:","ring-number:","ring-type:","spore-print-color:","population:","habitat:") 
names(newdata) <- gsub(":", "",names(newdata))
names(newdata) <- gsub("-", "_",names(newdata))
names(newdata)
head(newdata)
str(newdata)
summary(newdata)   
```


## Section 2: Change levels for only 3 columns 
```{r}
#check levels and change level Classes
levels(newdata$classes)[levels(newdata$classes)=="e"] <- "edible"
levels(newdata$classes)[levels(newdata$classes)=="p"] <- "poisonous"
#check levels and change level cap_shape
levels(newdata$cap_shape)[levels(newdata$cap_shape)=="b"] <- "bell"
levels(newdata$cap_shape)[levels(newdata$cap_shape)=="c"] <- "conical"
levels(newdata$cap_shape)[levels(newdata$cap_shape)=="f"] <- "flat"
levels(newdata$cap_shape)[levels(newdata$cap_shape)=="k"] <- "knobbed"
levels(newdata$cap_shape)[levels(newdata$cap_shape)=="s"] <- "sunken"
levels(newdata$cap_shape)[levels(newdata$cap_shape)=="x"] <- "convex"
#check levels and change level cap_surface
levels(newdata$cap_surface)[levels(newdata$cap_surface)=="f"] <- "fibrous"
levels(newdata$cap_surface)[levels(newdata$cap_surface)=="g"] <- "grooves"
levels(newdata$cap_surface)[levels(newdata$cap_surface)=="y"] <- "scaly"
levels(newdata$cap_surface)[levels(newdata$cap_surface)=="s"] <- "smooth"

for (n in names(newdata))
  if (is.factor(newdata[[n]])) {
    print (n)
      print(levels(newdata[[n]]))
  }

```

## Section 3: Use Tibble to change few more cols since changing levels using index was taking a while
```{r}
dftibble <- tibble::as.tibble(newdata)

#cols4
dftibble <- dftibble %>% 					
  mutate(cap_color = case_when(
    .$cap_color == "n" ~ "brown",
    .$cap_color == "b" ~ "buff",
    .$cap_color == "c" ~ "cinnamon",
    .$cap_color == "g" ~ "gray",
    .$cap_color == "p" ~ "pink",
    .$cap_color == "u" ~ "purple",
    .$cap_color == "e" ~ "red",
    .$cap_color == "w" ~ "white",
    .$cap_color == "y" ~ "yellow",
    TRUE ~ "other"))
#col5
dftibble <- dftibble %>% mutate(bruises = case_when(
  .$bruises == "t" ~ "bruises",
  .$bruises == "f" ~ "no"
  ,TRUE ~ "other"))
#col12 
dftibble <- dftibble %>% 
  mutate(stalk_root = case_when(
    .$stalk_root == "b" ~ "bulbous",
    .$stalk_root == "c" ~ "club",
    .$stalk_root == "u" ~ "cup",
    .$stalk_root == "e" ~ "equal",
    .$stalk_root == "z" ~ "rhiomorphus",
    .$stalk_root == "r" ~ "rooted",
    .$stalk_root == "?" ~ "missing"
    ,TRUE ~ "other"))
```

## Section 3: Convert tibble back to dataframe and report summary
```{r}
#convert to factor
dftibble <- mutate_at(dftibble, vars(cap_color, bruises,stalk_root), as.factor)
head(dftibble)
#convert back to dataframe
finaldata <- data.frame(dftibble)
nrow(finaldata)
str(finaldata)
```

## Section 4: create subset with those missing and non missing on stalk_root
```{r}
#Use file with StringasFactors=FALSE
allnonmissing <- finaldata[!finaldata$stalk_root == "missing",c(1,2:4,12) ] #subset rows with no missing for attribute
Submissingrows <- finaldata[finaldata$stalk_root == "missing",c(1,2:4,12) ] #subset only missing rows 
dim(allnonmissing)
dim(Submissingrows)
head(allnonmissing)
head(Submissingrows)
```
## Section 5: create subset with some conditions on rows
```{r}
df1 <- subset(allnonmissing, cap_color == "yellow" & cap_surface == "scaly" )
df2 <- subset(allnonmissing, cap_color %in% c("yellow","brown") & cap_surface == "scaly" )
summary(df1)
str(df1)
head(df1)
str(df2)
summary(df2)
head(df2)
```

